Project 4

### 2018-11-17 Project4/report/ms.tex
Updates from ShareLaTeX

\section{Method}
\label{sec:method}
h

The two-dimensional binary Ising model has analytical expectation values.

\begin{deluxetable}{cccc}
\tablewidth{0pt}
\tablecaption{\label{tab:results}}
\tablecomments{....}
\tablecolumns{4}
\tablehead{Spins up  & Degeneracy & Energy & Magnetic moment}
\startdata
4 & 1 & -8J & 4 \\
3 & 4 & 0 & 2 \\
2 & 4 & 0 & 0 \\
2 & 2 & 8J & 0 \\
1 & 4 & 0 & -2 \\
0 & 1 & -8J & -4
\enddata
\end{deluxetable}

>>>>

\section{Method}
\label{sec:method}
The two-dimensional binary Ising model has analytical expectation values. The partition function
%
\begin{equation*}
    z = \sum_i \Omega(E_i) e^{\beta E_i},
\end{equation*}
%
where $\Omega(E_i)$ is the degeneracy of energy $E_i$ and $\beta = 1/k_BT$. From Table \ref{tab:energies} we can see that there are 2 energies that are non-zero: $\pm 8J$, which both has a degeneracy of 2. The energy equal to zero has a degeneracy of 12. Expanding the sum of the partition function we get 
%
\begin{align*}
    z &= 2e^{8J\beta} + 12 + 2e^{-8J\beta} \\
    &= 12 + 2\left(e^{8J\beta} + e^{-8J\beta}\right),
\end{align*}
%
where we have inserted $e^0 = 1$. We know that $\cosh(x) = \frac{1}{2}\left( e^{-x} + e^x\right)$, inserting this gives
%
\begin{gather}\label{eq:z}
    Z = 12 + 4\cosh(8J\beta).
\end{gather}
%
The mean value of the energy is given by
%
\begin{align*}
    \langle E \rangle = -\frac{\partial }{\partial \beta} \ln Z = -\frac{1}{Z}\frac{\partial Z}{\partial \beta},
\end{align*}
%
which is found by inserting the partition function.
%
\begin{align*}
    \langle E \rangle &= -\frac{1}{Z}\frac{\partial}{\partial \beta} \left(12 + 4 \cosh\left(8J\beta\right)\right)\\
    &= -\frac{4}{Z}\left(8J\sinh\left(8J\beta\right)\right),
\end{align*}
pulling a factor 4 from the partition function and inserting we are left with a analytical solution to the mean energy of the lattice
%
\begin{gather*}
    \langle E \rangle = -\frac{8J\sinh\left(8J\beta\right)}{3 + \cosh(8J\beta}.
\end{gather*}
%
The heat capacity can be found by
%
\begin{align*}
    C_V &= \frac{1}{kT^2}\frac{\partial^2}{\partial \beta^2} \ln Z = \frac{1}{kT^2}\frac{\partial}{\partial \beta}\left(\frac{\partial}{\partial \beta} \ln Z\right),
\end{align*}
%
where we recognize the partial derivative in the parenthesis as the negative mean energy. 
%
\begin{align*}
    C_V = \frac{1}{kT^2}&\left(\frac{64J^2\cosh\left(8J\beta\right)\left(3 + \cosh\left(8J\beta\right)\right)}{\left(3 + \cosh\left(8J\beta\right)\right)^2}\right \\
    &-\left \frac{64J^2\sinh^2\left(8J\beta\right)}{\left(3 + \cosh\left(8J\beta\right)\right)^2}\right),
\end{align*}
%
by inserting for the mean energy and using the product rule. Further simplifications gives the mean heat capacity
%
\begin{gather*}
    C_V = \frac{1}{kT^2}\frac{64\cosh\left(8J\beta\right)}{ 3 + \cosh\left(8J\beta\right)}.
\end{gather*}
The susceptibility is given as
%
\begin{equation*}
    \chi = \frac{1}{kT}\left(\langle M^2 \rangle - \langle M \rangle^2\right).
\end{equation*}
%
To find the susceptibility of the system we consider the mean magnetic moment
%
\begin{equation*}
    \langle M \rangle = \frac{1}{Z} \sum_i |M_i| e^{-\beta E_i},
\end{equation*}
where we find the magnetic moment for a given energy in Table \ref{tab:energies}. So the mean magnetic moment is
%
\begin{equation*}
    \langle M \rangle = \frac{1}{Z}\left(4e^{8J\beta} + 4\left(2e^{0}\right) + 4\left(-2e^0\right) + -4e^{8J\beta} \right) = 0.
\end{equation*}
%
Further we calculate the mean of the  magnetic moment squared
%
\begin{equation*}
    \langle M^2 \rangle = \frac{1}{Z}\sum_i M_i^2 e^{-\beta E_i},
\end{equation*}
and again referring to Table \ref{tab:energies} we get
%
\begin{align*}
    \langle M^2 \rangle &= \frac{1}{Z}\left(16\left(2e^{8J\beta}\right) + 16\left(2e^0\right) \right) \\
    &= \frac{8\left(e^{8J\beta} + 1\right)}{3 + \cosh\left(8J\beta\right)}.
\end{align*}
So the susceptibility of the system is given as
%
\begin{equation*}
    \chi = \frac{1}{kT^2}\langle M^2 \rangle = \frac{1}{kT^2}\frac{8\left(e^{8J\beta} + 1\right)}{3 + \cosh\left(8J\beta\right)}
\end{equation*}
%
Further we find the mean of the absolute value of the magnetic moment by

\begin{align*}
    \langle |M| \rangle &= \frac{1}{Z} \sum_i |M_i| e^{-\beta E_i} \\
     &= \frac{1}{Z}\left(4e^{8J\beta} + 4\left(2e^{0}\right) + 4\left(|-2|e^0\right) + |-4|e^{8J\beta} \right) \\
     &= \frac{2e^{8J\beta} + 4}{3 + \cosh\left(8J\beta\right)},
\end{align*}
%
after pulling out a factor 4 from the partition function. To summarize we have the following analytical expression for the system
%
\begin{align*}
    \langle E \rangle &= -\frac{4}{Z}\left(8J\sinh\left(8J\beta\right)\right), \\
    \langle |M| \rangle &= \frac{2e^{8J\beta} + 4}{3 + \cosh\left(8J\beta\right)}, \\
    C_V &= \frac{1}{kT^2}\frac{64\cosh\left(8J\beta\right)}{ 3 + \cosh\left(8J\beta\right)}, \\
    \chi &= \frac{1}{kT^2}\frac{8\left(e^{8J\beta} + 1\right)}{3 + \cosh\left(8J\beta\right)}.
\end{align*}

\begin{deluxetable}{cccc}
\tablewidth{0pt}
\tablecaption{\label{tab:energies}}
\tablecomments{...}
\tablecolumns{4}
\tablehead{Spins up  & Degeneracy & Energy & Magnetic moment}
\startdata
4 & 1 & -8J & 4 \\
3 & 4 & 0 & 2 \\
2 & 4 & 0 & 0 \\
2 & 2 & 8J & 0 \\
1 & 4 & 0 & -2 \\
0 & 1 & -8J & -4
\enddata
\end{deluxetable}
